---
title: "tidytuesday: r4ds time series"
author: "Zach Bogart"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(scales)
```

```{r message=FALSE}
members <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-16/r4ds_members.csv")

```

```{r}
glimpse(members)
```

## Membership over time

Taking a look at total membership and how it has grown over time
```{r}
members %>%
  ggplot() +
  geom_line(aes(x=date, y=total_membership), color='red') +
```

There doesn't seem to be a difference between *full_members* and *total_membership*.
```{r}
members %>%
  select(c(date, total_membership, full_members)) %>%
  gather("member_type", "count", -date) %>%
  ggplot(aes(x=date, y=count, col=member_type)) +
  geom_line() +
  ggtitle("There is only one line...maybe these are the same columns?")
```

To confirm, there doesn't seem to be any difference between the two columns:
```{r}
members %>%
  mutate(member_diff = abs(full_members - total_membership)) %>%
  select(member_diff) %>%
  summarise(total_diff = sum(member_diff))
```

Would remove one of the columns.

## Active members, daily and weekly

Let's look at how the the r4ds active members changed over the time period.

The daily is hovering around 100-200 people, weekly can be more sporadic with spikes in the 400s.
```{r}
members %>%
  select(c(date, daily_active_members, weekly_active_members)) %>%
  gather("member_type", "count", -date) %>%
  ggplot(aes(x=date, y=count, col=member_type)) +
  geom_line() + 
  scale_x_date(breaks = pretty_breaks(22)) +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))
```

Covers about 22 months 
```{r}
date_range <- interval(first(members$date), last(members$date))
date_range / months(1)
```