---
title: "us_tuition_costs"
author: "Zach"
date: "1/29/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

## 2018 Week 1 (2018-04-02) US Tuition Costs

*tidytuesday*

```{r}
library(tidyverse)
library(readxl)
library(scales)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

theme_set(theme_light())
```

```{r}
file_path = "../data/us_avg_tuition.xlsx"

tuition_costs <- read_excel(file_path)
```

### Look at average tuition rates across states

```{r}
tuition_costs %>% 
  gather(year_range, tuition, -State) %>%
  group_by(State) %>%
  summarise(avg_tuition = mean(tuition)) %>%
  arrange(desc(avg_tuition))
```

Found Vermont has the largest average tuition rate.

### How Does Tuition Change in Vermont?

```{r}
tuition_costs %>%
  filter(State == "Vermont") %>%
  gather(year_range, tuition, -State) %>%
  arrange(State) %>% 
  ggplot(aes(year_range, tuition)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates Over Time for California",
       subtitle = "via onlinembapage.com") +
  coord_flip() +
  geom_point() 
```

### Throw Everything on one graph!

```{r}
tuition_costs %>%
  gather(year_range, tuition, -State) %>%
  arrange(State) %>% 
  ggplot(aes(year_range, tuition)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates Over All States",
       subtitle = "via onlinembapage.com") +
  coord_flip() +
  geom_point()
```

### Let's make some regions

Regions from https://en.wikipedia.org/wiki/List_of_regions_of_the_United_States. I feel like these regions could be hot-button subjects...

```{r}
# Northeast
northeast <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")
# Midwest
midwest <- c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota")
# South
south <- c("Delaware", "Florida", "Georgia", "Maryland", "North Carolina", "South Carolina", "Virginia", "West Virginia", "Alabama", "Kentucky", "Mississippi",  "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas")
# West
west <- c("Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington")

```

Make a version of the dataframe that has regions included.

```{r}
tuition_regions = tuition_costs %>%
  mutate(region = if_else(State %in% northeast, "northeast", 
                 if_else(State %in% midwest, "midwest",
                 if_else(State %in% south, "south", "west"))))
```

### Overlap Regions

Watch the gather when copying code over (can't forget to exclude the new region column or we run into trouble).

```{r}
tuition_regions %>%
  gather(year_range, tuition, -State, -region) %>% 
  ggplot(aes(year_range, tuition, color = region)) +
  scale_y_continuous(labels = dollar_format()) +
  geom_point(size = 4, alpha = 0.7) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates by Region",
       subtitle = "via onlinembapage.com") +
  coord_flip()
```

## Appendix

### Northeast

```{r}
tuition_costs %>%
  filter(State %in% northeast) %>%
  gather(year_range, tuition, -State) %>%
  arrange(State) %>% 
  ggplot(aes(year_range, tuition)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates in the Northeast",
       subtitle = "via onlinembapage.com") +
  coord_flip() +
  geom_point() +
  facet_wrap(~State)
```

### Midwest

```{r}
tuition_costs %>%
  filter(State %in% midwest) %>%
  gather(year_range, tuition, -State) %>%
  arrange(State) %>% 
  ggplot(aes(year_range, tuition)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates in the Midwest",
       subtitle = "via onlinembapage.com") +
  coord_flip() +
  geom_point() +
  facet_wrap(~State)
```

### South

```{r}
tuition_costs %>%
  filter(State %in% south) %>%
  gather(year_range, tuition, -State) %>%
  arrange(State) %>% 
  ggplot(aes(year_range, tuition)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates in the South",
       subtitle = "via onlinembapage.com") +
  coord_flip() +
  geom_point() +
  facet_wrap(~State)
```

### West

```{r}
tuition_costs %>%
  filter(State %in% west) %>%
  gather(year_range, tuition, -State) %>%
  arrange(State) %>% 
  ggplot(aes(year_range, tuition)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Year Range", y = "Tuition",
       title = "Tuition Rates in the West",
       subtitle = "via onlinembapage.com") +
  coord_flip() +
  geom_point() +
  facet_wrap(~State)
```