---
title: "Video Games Dataset"
author: "Zach"
date: "8/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Background

https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-07-30

## Let's Play!

```{r}
library(tidyverse)
```

```{r}
# clean dataset from lizawood's github
url <- "https://raw.githubusercontent.com/lizawood/apps-and-games/master/PC_Games/PCgames_2004_2018_raw.csv"

# read in raw data
raw_df <- url %>% 
  read_csv() %>% 
  janitor::clean_names() 

# clean up some of the factors and playtime data
clean_df <- raw_df %>% 
  mutate(price = as.numeric(price),
         score_rank = word(score_rank_userscore_metascore, 1),
         average_playtime = word(playtime_median, 1),
         median_playtime = word(playtime_median, 2),
         median_playtime = str_remove(median_playtime, "\\("),
         median_playtime = str_remove(median_playtime, "\\)"),
         average_playtime = 60 * as.numeric(str_sub(average_playtime, 1, 2)) +
           as.numeric(str_sub(average_playtime, 4, 5)),
         median_playtime = 60 * as.numeric(str_sub(median_playtime, 1, 2)) +
           as.numeric(str_sub(median_playtime, 4, 5)),
         metascore = as.double(str_sub(score_rank_userscore_metascore, start = -4, end = -3))) %>% 
  select(-score_rank_userscore_metascore, -score_rank, -playtime_median) %>% 
  rename(publisher = publisher_s, developer = developer_s)
```

## Explore

Let's see how price is spread out over the dataset.

```{r}
clean_df %>%
  ggplot(aes(x=price)) +
  geom_histogram()
```

Looks like there are some "video games" that are super expensive. But there aren't that many.

```{r}
clean_df %>%
  select(game, price) %>%
  filter(price > 100) %>%
  summarise(count=n())
```

Let's narrow in on where most of the data is.

```{r}
clean_df %>%
  select(game, price) %>%
  filter(price < 50) %>%
  ggplot(aes(price)) +
  geom_histogram()
```

There are spikes around the even-numbered price points (like 19.99 or 9.99).

## How much do different developers sell their games for?

```{r}
clean_df %>%
  na.omit() %>%
  select(price, game, developer) %>%
  group_by(developer) %>%
  summarise(average_cost = mean(price), game_count = n())
```

Looks like a lot of these developers only have put out one or two games. Maybe we can look at the publishers.

```{r}
game_costs <- clean_df %>%
  na.omit() %>%
  select(price, game, publisher) %>%
  group_by(publisher) %>%
  summarise(average_cost = mean(price), game_count = n()) %>%
  arrange(desc(game_count))
```

```{r}
game_costs %>%
  head()
```

