---
title: "Roman Emperors"
author: "Zach"
date: "8/13/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
emperors <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-13/emperors.csv")
```

## Start

Ideas:
- Plot a line graph of the birth and death dates of the emperors
- Group by birth city
- Look into rise variable
- Look at breakdown of cause of death

## Explore
```{r}
glimpse(emperors)
```

## Cause of death

```{r}
emperors %>%
  group_by(cause) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

The majority of the emperors were either killed by assassination or natural causes. 

```{r}
emperors %>%
  group_by(cause) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ggplot(aes(reorder(cause, count), count)) +
    geom_col() + 
    labs(x="Cause of Death", y="Death Count", title="How They Died") +
    coord_flip()
```

## Lifespan

```{r}
emperors %>%
  filter(!is.na(birth), !is.na(death)) %>%
  gather(key = "lifespan", value = "date", birth, death, -name) %>%
  ggplot(aes(date, reorder(name, desc(date)), color = lifespan)) +
    geom_point() +
    labs(x="Year", y="Emperor", title="When Emperors Were Alive")
```

We can see when the emperors were alive. We also see that a couple of the emperors died before they were born! Guessing that is a typo and would look into swapping the values. Moving on.

Looking at the overlap of lifespans for the different emperors, there are many overlapping rulers. Let's do the same thing, but looking at just their reign.

```{r}
emperors %>%
  filter(!is.na(birth), !is.na(death)) %>%
  gather(key = "reign_span", value = "date", reign_start, reign_end, -name) %>%
  ggplot(aes(date, reorder(name, desc(date)), color = reign_span)) +
    geom_point() +
    labs(x="Year", y="Emperor", title="When Emperors Ruled")
```

We see it is much tighter with much less overlap between rulers. Their reigns were quite much shorter. Look at some averages of the reigns and lifespans.

```{r}
emperors_life <- emperors %>%
  filter(!is.na(birth), !is.na(death)) %>%
  mutate(lifespan = death - birth,
         reign_span = reign_end - reign_start)

cat("Average Lifespan:", mean(emperors_life$lifespan))
cat("Average Reign:", mean(emperors_life$reign_span))
```

The emperors were living, on average, to the ripe old age of `r round(mean(emperors_life$lifespan) / 365)`, ruling around `r round(mean(emperors_life$reign_span) / 365)` years. 

## Rise

```{r}
emperors %>%
  group_by(rise) %>%
  summarise("count"=n()) %>%
  arrange(desc(count))
```

Looks like the most popular way to become emperor was to be born into it. Democratic Elections weren't really in vogue at the time.

## Birth City

The cities look pretty specific, so let's get a broad sense of where they are from with the `birth_prv` column.

```{r}
emperors %>%
  group_by(birth_prv) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

There are a bunch of unknowns, but a lot of the emperors in this dataset are from Italy and [Pannonia](https://en.wikipedia.org/wiki/Pannonia), a province of the Roman Empire. So many of the Roman Emperors were born in Italy, but a bunch, of course, were also born within the Roman Empire itself.